Model Finding for Recursive Functions in SMT

### 2016-01-26 ijcar2016/body.tex
more edits

Nunchaku parses and types the input problem before applying a sequence
of translations, each reducing the distance to the target fragment.
In our example, the predicates
\verb!even! and \verb!odd! are \emph{polarized} (specialized into a pair of
predicates such that one is used in positive positions and the other in
negative positions),
$m$ is skolemized into a fresh constants,
then translated into admissible recursive functions,
before another pass applies the encoding described in this paper.
If a model is found, it is translated back to the input language,
with \verb$?__$ placeholders as necessary.

Conceptually, the sequence of transformation is a \relax{two-way pipeline}
built by composing pairs $(\textsf{encode}, \textsf{decode})$ of
transformations. For each such pair, \textsf{encode} transforms a problem over
a signature $\Sigma$ in a logic $\mathcal{L}$ to a problem over a signature
$\Sigma'$ in a logic $\mathcal{L}'$, and \textsf{decode} translates a model in
$\mathcal{L'}$ over $\Sigma'$ into a model in $\mathcal{L}$ over $\Sigma$,
in the spirit of institution theory \cite{goguen-burstall-1992}. The
pipeline currently implemented in Nunchaku consists of the following phases:

\begin{description}
%{{{
  \item[Type inference] infers types and checks definitions;
  \item[Type skolemization]
    replaces $\exists \alpha.\; \varphi[\alpha]$
      with $\varphi[\tau]$, where $\tau$ is a fresh type;
  \item[Monomorphization]
    specializes polymorphic definitions on their type arguments and removes
    unused definitions;
  \item[Elimination of equations]
    translates multiple-equation definitions of recursive functions into
      a single nested pattern matching;
  \item[Polarization]
    specializes predicates into a version used in positive positions
      and a version used in negative positions;
  \item[Unrolling]
    adds a decreasing argument to possibly ill-founded predicates to make
    them well founded;
  \item[Skolemization]
    introduces Skolem symbols for term variables;
  \item[Elimination of Predicates]
    changes a multiple-clauses (co)inductive predicate definition
      into a single recursive function;
  \item[Recursion elimination]
    performs the encoding from Section~\ref{sec:encoding};
  \item[Elimination of pattern matching]
    rewrites pattern-matching expressions using datatype
    discriminators and selectors;
  \item[CVC4 invocation] runs CVC4 to obtain a model.
%}}}
\end{description}

>>>>

Nunchaku parses and types the input problem before applying a sequence
of translations, each reducing the distance to the target fragment.
In our example, the predicates
\verb!even! and \verb!odd! are \emph{polarized} (specialized into a pair of
predicates such that one is used in positive positions and the other in
negative positions),
$m$ is skolemized into a fresh constants,
then translated into admissible recursive functions,
before another pass applies the encoding described in this paper.
If a model is found, it is translated back to the input language,
with \verb$?__$ placeholders indicating unknown values.

Conceptually, the sequence of transformation is a \relax{two-way pipeline}
built by composing pairs $(\textsf{encode},\vthinspace \textsf{decode})$ of
transformations. For each such pair, \textsf{encode} transforms a problem over
a signature $\Sigma$ in a logic~$\mathcal{L}$ to a problem over a signature
$\Sigma'$ in a logic~$\mathcal{L}'$, and \textsf{decode} translates a model
in~$\mathcal{L}'$ over $\Sigma'$ into a model in~$\mathcal{L}$ over $\Sigma$,
in the spirit of institution theory \cite{goguen-burstall-1992}. The
pipeline currently consists of the following phases:

\begin{description}
%{{{
  \item[Type inference] infers types and checks definitions;
  \item[Type skolemization]
    replaces $\exists \alpha.\; \varphi[\alpha]$
      with $\varphi[\tau]$, where $\tau$ is a fresh type;
  \item[Monomorphization]
    specializes polymorphic definitions on their type arguments and removes
    unused definitions;
  \item[Elimination of equations]
    translates multiple-equation definitions of recursive functions into
      a single nested pattern matching;
  \item[Polarization]
    specializes predicates into a version used in positive positions
      and a version used in negative positions;
  \item[Unrolling]
    adds a decreasing argument to possibly ill-founded predicates to make
    them well founded;
  \item[Skolemization]
    introduces Skolem symbols for term variables;
  \item[Elimination of (co)inductive predicates]
    recasts a multiple-clause (co)inductive predicate specification
      into a recursive equation;
  \item[Recursion elimination]
    performs the encoding from Section~\ref{sec:encoding};
  \item[Elimination of pattern matching]
    rewrites pattern-matching expressions using datatype
    discriminators and selectors;
  \item[CVC4 invocation] runs CVC4 to obtain a model.
%}}}
\end{description}
